---
title: "ctlars Algorithm"
output:
  html_document:
    keep_md: true
  md_document:
    variant: gfm
  pdf_document: default
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
# Store user's options()
old_options <- options()

if (!dir.exists("man/figures/"))
{
  dir.create("man/figures/")
}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ctlars
**Title**: The CTLARS Algorithm: Complex early-terminated Forward Variable Selection

**Description**: It computes the solution path of the complex Terminating-LARS (CTLARS) algorithm.
The CTLARS algorithm appends dummy predictors to the original predictor matrix and terminates the forward-selection process after a pre-defined number of dummy variables has been selected.

In the following, we show how to use the package and give you an idea of why terminating the solution path early is a reasonable approach in high-dimensional and sparse variable selection: In many applications, most active variables enter the solution path early!
  

# Installation

You can install the development version of ctlars from [GitHub](https://github.com/G4Lactus/CTLARS_Package) with:

``` r
# install.packages("devtools")
# devtools::install_github("G4Lactus/CTLARS_Package")
```

# Quick Start
In the following, we illustrate the basic usage of the `ctlars` package for
performing variable selection in space high-dimensional complex data space,
using the CTLARS algorithm.

1. **First**, we generate high-dimensional Gaussian data set with sparse support.
```{r, echo=TRUE, fig.align='center', message=TRUE, fig.width=10, fig.height=6, out.width="90%"}
library(ctlars)

generate_simdata <- FALSE
if (generate_simdata) {
  n <- 100
  p <- 150
  beta_cardinality <- 5
  set.seed(42)

  # generate some data
  data <- generate_ccg_data(n,
                            p,
                            mean_xr = 0,
                            sd_xr = 1,
                            mean_xi = 0,
                            sd_xi = 1,
                            beta_cardinality = beta_cardinality,
                            noise_power = 1,
                            set_snr = TRUE,
                            snr_is_linear = TRUE,
                            snr_val_linear = 2,
                            snr_val_db = 10)

} else {
  data("WCCCGauss_data")
  data <- WCCCGauss_data
}
```


2. **Second**, we generate a dummy matrix containing `n` rows and `num_dummies`
 dummy predictors sampled from a white circularly centered symmetric complex
 Gaussian and append it to the original predictor matrix.

```{r, echo=TRUE, fig.align='center', message=TRUE, fig.width=10, fig.height=6, out.width="90%"}
 # Append dummies
 set.seed(789)
 update_data <- function(data_lst, num_dummies = p) {
   p <- ncol(data_lst$X)
   n <- nrow(data_lst$X)
 
   data_lst$X <- cbind(data_lst$X,
     matrix(complex(real = stats::rnorm(n * num_dummies, sd = 1),
                    imag = stats::rnorm(n * num_dummies, sd = 1)),
            nrow = n, ncol = num_dummies)
   )
   
   data_lst$num_dummies <- num_dummies
   data_lst$n <- n
   data_lst$p <- p
 
   return(data_lst)
 }
 data <- update_data(data, num_dummies = ncol(data$X))
```
 

3. **Third**, we generate an object of class `ctlars` and supply the information:
```{r, echo=TRUE, fig.align='center', message=TRUE, fig.width=10, fig.height=6, out.width="90%"}
# Run complex terminating lars algorithm
# ----------------------------------------
ctlars_obj <- ctlars::ctlars$new(data$X,
                                  data$y,
                                  num_dummies = data$num_dummies
                                  )
``` 

4. **Fourth**, we perform on CTLARS step on `ctlars_obj`, i.e., the CTLARS
algorithm is run until **t_stop = 1** dummy has entered the solution path
and stops there:
```{r terminated_solution_t_1, echo=TRUE, fig.align='center', message=TRUE, fig.width = 10, fig.height = 6, out.width = "90%"}
ctlars_obj$execute_clars_step(t_stop = 1, early_stop = TRUE, use_chol = TRUE)
```

5. **Five** We evaluate the selection results after **t_stop = 1** dummies have
been selected.
```{r, echo=TRUE, fig.align='center', message=TRUE, fig.width=10, fig.height=6, out.width="90%"}
# Test
# ----------------------------------------
cat("Proposed index set by CLARS: ", ctlars_obj$get_active_set(), "\n")
cat("True support: ", drop(data$support), "\n")
cat("Proposal in true support: ",
     intersect(ctlars_obj$get_active_set(), drop(data$support)), "\n")
cat("Dummies in proposal: ", ctlars_obj$get_active_dummies(), "\n")
 
plot(ctlars_obj)
```
