---
title: "The Complex Terminating-LARS (T-LARS) Method: Usage and Simulations"
author: |
  | Fabian Scheidt^\#^, Jasin Machkour ^\#^, Michael Muma^\#^
  |
  | ^\#^Technische Universität Darmstadt
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: flatly
    highlight: pygments
    toc: yes
    toc_depth: 1
    toc_float: yes
    css: vignette_styles.css
    mathjax: "https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.5/MathJax.js?config=TeX-AMS_CHTML.js"
  prettydoc::html_pretty:
    theme: tactile
    highlight: vignette
    toc: yes
    toc_depth: 2
toc-title: "Table of Contents"    
csl: ieee.csl
bibliography: refs.bib
nocite: |
  
vignette: |
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{The Complex Terminating-LARS (CTLARS) Method: Usage and Simulations}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
# Store user's options()
old_options <- options()

library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.retina = 2,
  out.width = "85%",
  dpi = 96
  # pngquant = "--speed=1"
)
options(width = 80)
```

-----------

# Motivation
This package implements the **Complex Terminating-LARS (CTLARS)** algorithm.
It computes the solution path of the CTLARS algorithm.
The CTLARS algorithm is a variation of a complex LARS algorithm first described
in [@graff2020reduced] and appends dummy predictors to the original predictor
matrix and terminates the forward-selection process after a pre-defined number of
dummy variables has been selected. The idea was first introduced by its real
number space counterpart the **Terminating-LARS (T-LARS)** algorithm according
to [@machkour2022terminating] itself inspired from the original work on the
**Least Angle Regression** algorithm [@efron2004least].

The motivation is the same as with its cousing. The classical LARS algorithm
computes the entire solution path without stopping early.
However, for a variety of applications there exists very little or no useful
information in later steps or the solution path, and in later step usually
only non-active (null) variables are selected.
Stopping early benefits to decrease computation time, while the accuracy is not
reduced.

In the following, we show how to use the package and give an intuitive idea
of why terminating the solution path early is a reasonable approach in
high-dimensional complex valued data and spare variable selection.
The answer is simply, that in most applications active variables enter the
solution path earlier compared to nulls!

You can open the help pages with:

```{r, eval=FALSE, echo=TRUE}
library(ctlars)
help(package = ctlars)
?`ctlars-package`
?ctlars
?FDP
?generate_ccg_data
?plot.ctlars
?simulate_model_TPR_FDR
?TPP
?WCCCGauss_data
?WCCCGauss_data_small
```

To cite the package `ctlars` in publications use
```{r, eval=FALSE}
# some future citation:
citation("ctlars")
```

<!-- Include Latex macros -->
$$
\DeclareMathOperator{\FDP}{FDP}
\DeclareMathOperator{\FDR}{FDR}
\DeclareMathOperator{\TPP}{TPP}
\DeclareMathOperator{\TPR}{TPR}
\newcommand{\A}{\mathcal{A}}
\newcommand{\coloneqq}{\mathrel{\vcenter{:}}=}
$$

# Installation

Currently no public installation is available.
``` r
install.packages("devtools")
devtools::install_github("to/be/announced")
```
******

# Quick Start
In the following, we illustrate the basic usage of the `ctlars` package for
performing variable selection in space high-dimensional complex data space,
using the CTLARS algorithm.

1. **First**, we generate high-dimensional Gaussian data set with sparse support.
```{r, echo=TRUE, fig.align='center', message=TRUE, fig.width=10, fig.height=6, out.width="90%"}
library(ctlars)

generate_simdata <- FALSE
if (generate_simdata) {
  n <- 100
  p <- 150
  beta_cardinality <- 5
  set.seed(42)

  # generate some data
  data <- generate_ccg_data(n,
                            p,
                            mean_xr = 0,
                            sd_xr = 1,
                            mean_xi = 0,
                            sd_xi = 1,
                            beta_cardinality = beta_cardinality,
                            noise_power = 1,
                            set_snr = TRUE,
                            snr_is_linear = TRUE,
                            snr_val_linear = 2,
                            snr_val_db = 10)

} else {
  data("WCCCGauss_data")
  data <- WCCCGauss_data
}
```


2. **Second**, we generate a dummy matrix containing `n` rows and `num_dummies`
 dummy predictors sampled from a white circularly centered symmetric complex
 Gaussian and append it to the original predictor matrix.

```{r, echo=TRUE, fig.align='center', message=TRUE, fig.width=10, fig.height=6, out.width="90%"}
 # Append dummies
 set.seed(789)
 update_data <- function(data_lst, num_dummies = p) {
   p <- ncol(data_lst$X)
   n <- nrow(data_lst$X)
 
   data_lst$X <- cbind(data_lst$X,
     matrix(complex(real = stats::rnorm(n * num_dummies, sd = 1),
                    imag = stats::rnorm(n * num_dummies, sd = 1)),
            nrow = n, ncol = num_dummies)
   )
   
   data_lst$num_dummies <- num_dummies
   data_lst$n <- n
   data_lst$p <- p
 
   return(data_lst)
 }
 data <- update_data(data, num_dummies = ncol(data$X))
```
 

3. **Third**, we generate an object of class `ctlars` and supply the information:
```{r, echo=TRUE, fig.align='center', message=TRUE, fig.width=10, fig.height=6, out.width="90%"}
# Run complex terminating lars algorithm
# ----------------------------------------
ctlars_obj <- ctlars::ctlars$new(data$X,
                                  data$y,
                                  num_dummies = data$num_dummies
                                  )
``` 

4. **Fourth**, we perform on CTLARS step on `ctlars_obj`, i.e., the CTLARS
algorithm is run until **t_stop = 1** dummy has entered the solution path
and stops there:
```{r terminated_solution_t_1, echo=TRUE, fig.align='center', message=TRUE, fig.width = 10, fig.height = 6, out.width = "90%"}
ctlars_obj$execute_clars_step(t_stop = 1, early_stop = TRUE, use_chol = TRUE)
```

5. **Five** We evaluate the selection results after **t_stop = 1** dummies have
been selected.
```{r, echo=TRUE, fig.align='center', message=TRUE, fig.width=10, fig.height=6, out.width="90%"}
# Test
# ----------------------------------------
cat("Proposed index set by CLARS: ", ctlars_obj$get_active_set(), "\n")
cat("True support: ", drop(data$support), "\n")
cat("Proposal in true support: ",
     intersect(ctlars_obj$get_active_set(), drop(data$support)), "\n")
cat("Dummies in proposal: ", ctlars_obj$get_active_dummies(), "\n")
 
plot(ctlars_obj)
```

******

# CTLARS Warm Starts
The object `ctlars_obj` stores the state information and, therefore, allows
for warm starts. After performing a CTLARS step, e.g., `t_stop = 1`, we can
perform another CTLARS, e.g., with `t_stop = 5`. That means the algorithm
iterates now till `5` dummies have been selected.
Subsequently, we plot again the solution path.

* **T_stop = 5**:
```{r Warm_Start, echo=TRUE, fig.align='center', message=TRUE, fig.width=10, fig.height=6, out.width="90%"}
ctlars_obj$execute_clars_step(t_stop = 5)
plot(ctlars_obj)
```

Of course these sort of warm starts deliver the same result, as if we had
started from scratch until `t_stop = 5`.
```{r terminated_solution_t_5, echo=TRUE, fig.align='center', message=TRUE, fig.width=10, fig.height=6, out.width="90%"}
# Try a second example with t_stop = 5
# ---------------------------------------
ctlars_obj <- ctlars$new(data$X,
                         data$y,
                         has_intercept = FALSE,
                         num_dummies = data$num_dummies)

ctlars_obj$execute_clars_step(t_stop = 5, early_stop = TRUE)
plot(ctlars_obj)
```

******

# CTLARS Full Solution Path

The algorithm allows also to compute the whole solution path.

```{r, eval = TRUE}
ctlars_obj <- ctlars$new(data$X,
                          data$y,
                          has_intercept = FALSE,
                          num_dummies = data$num_dummies, 
                          verbose = FALSE)
 
ctlars_obj$execute_clars_step(t_stop = 1, early_stop = FALSE)

plot(ctlars_obj)
```

******

# FDR and TPR

From a statistical point of view, it is desirable to use a variable selection method that allows for controlling the expected value of the FDP at a user-defined target level $\alpha \in [0, 1]$ while maximizing the number of selected variables. These type of methods exist and are called false discovery rate (FDR)-controlling methods. 


**Definitions** (FDR and TPR)
Let $\widehat{\A} \subseteq \lbrace 1, \ldots, p \rbrace$ be the set of selected variables, $\A \subseteq \lbrace 1, \ldots, p \rbrace$ the set of true active variables, $| \widehat{\A} |$ the cardinality of $\widehat{\A}$, and define $1 \lor a \coloneqq \max\lbrace 1, a \rbrace$, $a \in \mathbb{R}$. Then, the false discovery rate (FDR) and the true positive rate (TPR) are defined by
$$
\FDR \coloneqq \mathbb{E} \big[ \FDP \big] \coloneqq \mathbb{E} \left[ \dfrac{\big| \widehat{\A} \backslash \A \big|}{1 \lor \big| \widehat{\A} \big|} \right]
$$

and 

$$
\TPR \coloneqq \mathbb{E} \big[ \TPP \big] \coloneqq \mathbb{E} \left[ \dfrac{| \A  \cap \widehat{\A} |}{1 \lor | \A |} \right],
$$

respectively.


```{r}
 # Evaluate FDP and TPP
 # ----------------------------------------
 chosen <- ctlars_obj$get_active_set()
 # remove dummies
 chosen <- setdiff(
   chosen, ctlars_obj$get_active_dummies()
 )
 FDP(chosen, drop(data$support))
 TPP(chosen, drop(data$support))
 
 # Plot R2 and SSR
 # ------------------
 plot(ctlars_obj$get_ssr(), type = "l", xlab = "Lars Step", ylab = "SSR")
 grid()
 
 plot(ctlars_obj$get_r2(), type = "l", xlab = "Lars Step", ylab = "R2")
 grid()
```
 
 ******
 
# Simulations
We conduct Monte Carlo simulations and plot the resulting averaged FDP and TPP over the number of included dummies T.
Note that the averaged FDP and TPP are estimates of the FDR and TPR, respectively.

```{r FDR_and_TPR_sim, eval=TRUE, echo=TRUE, fig.align='center', message=FALSE, fig.width = 10, fig.height = 5, out.width = "90%"}
library(ctlars)

t_vec <- c(1, 2, 5, 10, 20, 50, 100)
sim_res <- simulate_model_TPR_FDR(
   t_vec,
   MC = 500,
   n = 100,
   p = 300,
   num_actives = 10,
   num_dummies = 300,
   set_snr = FALSE,
   verbose = FALSE
   )

```

```{r FDR_and_TPR_plot, eval=TRUE, echo=FALSE, fig.align='center', message=FALSE, fig.width = 10, fig.height = 5, out.width = "90%"}
# Plot results
library(ggplot2)
library(patchwork)

# data frame containing data to be plotted (FDR and TPR in %)
# -------------------------------------------------------------
plot_data <- data.frame(T_vec = t_vec,
                        FDR = 100 * sim_res$FDR,
                        TPR = 100 * sim_res$TPR
                        )
                        
 
# FDR vs. T
# ------------
fdr_vs_t <- ggplot(plot_data, aes(x = t_vec, y = FDR)) +
   labs(x = "T", y = "FDR") +
   scale_x_continuous(breaks = t_vec[-2], minor_breaks = c(2),
                      limits = c(t_vec[1], t_vec[length(t_vec)])) +
   scale_y_continuous(breaks = seq(0, 100, by = 10),
                      minor_breaks = c(), limits = c(0, 100)) +
   geom_line(linewidth = 1.5, colour = "#336C68") +
   geom_point(size = 2.5, colour = "#336C68") +
   theme_bw(base_size = 16) +
   theme(
     panel.background = element_rect(fill = "white",
                                     color = "black",
                                     linewidth = 1)
   ) +
   coord_fixed(ratio =  0.85 * t_vec[length(t_vec)] / (100 - 0))


# TPR vs. T
# ------------
tpr_vs_t <- ggplot(plot_data, aes(x = t_vec, y = TPR)) +
   labs(x = "T", y = "TPR") +
   scale_x_continuous(breaks = t_vec[-2], minor_breaks = c(2),
                      limits = c(t_vec[1], t_vec[length(t_vec)])) +
   scale_y_continuous(breaks = seq(0, 100, by = 10),
                      minor_breaks = c(), limits = c(0, 100)) +
   geom_line(linewidth = 1.5, colour = "#336C68") +
   geom_point(size = 2.5, colour = "#336C68") +
   theme_bw(base_size = 16) +
   theme(
     panel.background = element_rect(fill = "white",
                                     color = "black",
                                     linewidth = 1)
   ) +
   coord_fixed(ratio =  0.85 * t_vec[length(t_vec)] / (100 - 0))
 
fdr_vs_t + tpr_vs_t
 
 
# TPR vs. FDR
# ------------
tpr_vs_fdr <- ggplot(plot_data, aes(x = FDR, y = TPR)) +
   labs(x = "FDR",
        y = "TPR") +
   geom_line(linewidth = 1.5, colour = "#336C68") +
   geom_point(size = 2.5, colour = "#336C68") +
   theme_bw(base_size = 16) +
   theme(panel.background = element_rect(fill = "white",
                                         color = "black",
                                         linewidth = 1)
         )

tpr_vs_fdr
 
```

As with the real T-LARS algorithm, we observe again that with growing number of included dummies
`t` the FDR and TPR increases.
Moreover, we can again identify the trade-off in the receiver-operating curve between FDR and TPR.

# Outlook
The CTLARS algorithm is going to be a major building block the complex T-Rex selector and its future variants.
Track its developement!

# References
<!-- <div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0px;"></div> -->
Currently nothing to quote.
